Models of lens candidates from Space Warps CFHTLS

### 2017-02-28 parts/intro.tex
edits
update intro, morph and table

With no unambiguous dark-matter particle detections so far, dark
matter has studied only through its indirect consequences on galactic
and larger scales.  In most work, dark matter is taken to be a
collisionless non-relativistic fluid (cold dark matter or CDM); in
recent years, the simulation by \cite{2005Natur.435..629S} has been
particularly influential in studying the formation of CDM structures.
Other scenarios have also been considered, such as a cold condensing
boson fluid \citep{2016ApJ...818...89S}.  There is a general
consensus, however, that the origins of galaxy lie in the
gravitational collapse, fragmentation, and mergers of dark-matter
clumps, into which fell gas, cooling through radiative processes to
form dense clouds and eventually stars.  There is, however, much
debate about the details, of which \cite{2012RAA....12..917S} provide
a nice summary.

@@

In this paper, we present SpaghettiLens models of lens-candidates
discovered through Space~Warps.  Purpose is twofold.  First, see if
plausible mass models can produce the overall configuration.  Second,
interpretation as lensing galaxies.

Each lens candidates is modelled in a collaborative refinement
process, where anyone interested can creata a new model or modify an
existing model to try and make it better.\footnote{This is in contrast
  to the main Space~Warps project for discovering lenses, where
  volunteers in a crowd of $\gtrsim10^4$ make independent
  contributions.  Each person is presented with a random selection of
  survey-patches and invited to (in effect) vote on each.  The system
  estimates each volunteer's skill level according to test-patches
  interspersed with the real data, and weights their votes accordingly
  \citep{2016MNRAS.455.1171M}.  There is an active forum for
  volunteers, but since everyone is seeing different data samples with
  minimal overlap, the forum has little if any influence on votes.}
The result represents a consensus among contributors, as to the best
that could be achieved with the available data and software.

@@

\begin{itemize}
\item First we have diagnostics based on morphology of the
  system.\todo{Make ordering consistent in text and
    Table~\ref{tab:models}.}  Section~\ref{sec:morph} and
  Figure~\ref{fig:splinput} explain.
\begin{itemize}
\item Whether the images are unblended.  Distinct unblended images are
  an advantage in modelling, but not essential.
\item Whether all images are discernable.  The topography of an
  arrival-time surface, as encoded by a spaghetti diagram, may require
  more images than are visible, in which case the modeller has to
  insert conjectural image positions.
\item Whether the lens is fairly isolated.
\item The image morphology concisely described: double or quads,
  further sub-categorised to indicate the elongation direction of the
  lensing mass.
\end{itemize}
\item Second we have mass models, covered in Section~\ref{sec:massmodels}.
\begin{itemize}
\item Whether the mass map is reasonable. Figure~\ref{fig:kappa}.
\item Whether the arrival-time surface and synthetic image are
  plausible.  For SW36 the model implies extra images or a long arc,
  which are not seen.  Figures~\ref{fig:arriv} and \ref{fig:synth} and
  \ref{fig:encl}.
\end{itemize}
\item Third, whether the implied lensing mass is plausible, given the
  star-light in the lensing galaxy.  Section~\ref{sec:stellar-mass}
  explains how we compare the lensing mass with the mass in stars in
  the lensing galaxy.  We estimate the stellar mass by comparing
  galaxy magnitudes from the CFHTLS pipeline with the well-known
  stellar-population models of \cite{2003MNRAS.344.1000B}.  We then
  extrapolate the stellar masses to a halo mass using the
  abundance-matching prescription of \cite{2010ApJ...710..903M}.
  Naturally, the lensing mass must be more than the stellar mass but
  no more than the total halo mass.  We then introduce what we call a
  halo index $\haloindex$ which gives an idea of how the lensing mass
  compares with these two bounds.  Figure~\ref{fig:stelmass}.
\end{itemize}

@@

The online supplement gives results for all the modelled systems.

>>>>

With no unambiguous dark-matter particle detections so far, dark
matter has studied only through its indirect consequences on galactic
and larger scales.  In most work, dark matter is taken to be a
collisionless non-relativistic fluid (cold dark matter or CDM); in
recent years, the ``millenium'' simulation by \cite{2005Natur.435..629S} has been
particularly influential in studying the formation of CDM structures.
Other scenarios have also been considered, such as a cold condensing
boson fluid \citep{2016ApJ...818...89S}.  There is a general
consensus, however, that the origins of galaxy lie in the
gravitational collapse, fragmentation, and mergers of dark-matter
clumps, into which fell gas, cooling through radiative processes to
form dense clouds and eventually stars.  There is, however, much
debate about the details, of which \cite{2012RAA....12..917S} provide
a nice summary.

@@

In this paper, we present SpaghettiLens models of lens-candidates
discovered through Space~Warps.  It's purpose is twofold: First, see if
plausible mass models can produce the overall configuration.  Second,
interpretation as lensing galaxies.

Each lens candidate is modelled in a collaborative refinement
process, where anyone interested can create a new model or modify an
existing model to try and make it better.\footnote{This is in contrast
  to the main Space~Warps project for discovering lenses, where
  volunteers in a crowd of $\gtrsim10^4$ make independent
  contributions.  Each person is presented with a random selection of
  survey-patches and invited to (in effect) vote on each.  The system
  estimates each volunteer's skill level according to test-patches
  interspersed with the real data, and weights their votes accordingly
  \citep{2016MNRAS.455.1171M}.  There is an active forum for
  volunteers, but since everyone is seeing different data samples with
  minimal overlap, the forum has little if any influence on votes.}
The result represents a consensus among contributors, as to the best
that could be achieved with the available data and software.

@@

\begin{itemize}
\item First we have diagnostics based on morphology of the
  system.
  Section~\ref{sec:morph} and Figure~\ref{fig:splinput} explain.
\begin{itemize}
\item The image morphology concisely described: double or quads,
  further sub-categorised to indicate the elongation direction of the
  lensing mass.
\item Whether the images are unblended.  Distinct unblended images are
  an advantage in modelling, but not essential.
\item Whether all images are discernable.  The topography of an
  arrival-time surface, as encoded by a spaghetti diagram, may require
  more images than are visible, in which case the modeller has to
  insert conjectural image positions.
\item Whether the lens is fairly isolated.
\end{itemize}
\item Second we have mass models, covered in Section~\ref{sec:massmodels}.
\begin{itemize}
\item Whether the mass map is reasonable. Figure~\ref{fig:kappa}.
\item Whether the arrival-time surface and synthetic image are
  plausible.  In particular, additional images are implied in regions
  where they are not observed signal an unsatisfactory model.
  Figures~\ref{fig:arriv} and \ref{fig:synth} and \ref{fig:encl}.
\end{itemize}
\item Third, whether the implied lensing mass is plausible, given the
  star-light in the lensing galaxy.  Section~\ref{sec:stellar-mass}
  explains how we compare the lensing mass with the mass in stars in
  the lensing galaxy.  We estimate the stellar mass by comparing
  galaxy magnitudes from the CFHTLS pipeline with the well-known
  stellar-population models of \cite{2003MNRAS.344.1000B}.  We then
  extrapolate the stellar mass to a halo mass using the
  abundance-matching prescription of \cite{2010ApJ...710..903M}.
  Naturally, the lensing mass must be more than the stellar mass but
  no more than the total halo mass.  We then introduce what we call a
  halo index $\haloindex$ which gives an idea of how the lensing mass
  compares with these two bounds.  Figure~\ref{fig:stelmass}.
\end{itemize}

@@

The online supplement gives results for all the modelled systems generated
for all the lensing candidates.

### 2017-02-28 parts/morph.tex
update intro, morph and table

We now discuss the diagnostics that can be given from the process of
drawing the spaghetti diagram, even before detailed mass-modelling
takes place.  Figure~\ref{fig:splinput} shows nine examples, each
consisting of a cutout of the Space~Warps image of a lens candidate,
marked up with a spaghetti diagrams.

@@

The fourth diagnostic is based on the fact that the arrangement of
lensed images of a pointlike source through a non-circular
gravitational lens depends on the location of the source relative to
the long and short axes of the lens.  This dependence is quite robust
and independent of many other details of the lensing mass
distribution.  Sources close to being dead-centre behind a lens tend
to produce quads, sources at larger transverse distance tend to
produce doubles.  We denote these as Q and D respectively, and add a
prefix to the Q systems, as follows: we write LQ if the source is
inferred as displaced along the long axis of the lens, SQ if displaced
along the short axis, IQ if inclined to both axes.  Although the
unlensed source and its location are obviously not seen, the LQ, SQ
and IQ cases correspond to easily-seen image morphologies \citep[see,
  e.g.,][]{2003AJ....125.2769S}.
\begin{itemize}
\item The simplest is the SQ case: this creates a saddle point and two
  minima in an arc, with the second saddle point on the other side of
  the lensing galaxy, and closer to the galaxy than the arc is.  SW58
  and SW28 in the upper row of Figure~\ref{fig:splinput} are typical
  examples of SQ.
\item The middle row of Figure~\ref{fig:splinput} shows three IQ
  systems, SW05, SW42 and SW19.  This type has a characteristic
  asymmetry, often with two images close together.
\item The lower row of Figure~\ref{fig:splinput} shows three LQ
  systems, SW09, SW29 and SW02, and the failed model SW57 at top right
  also possibly belonging to this category.  Here the images have a
  fairly symmetric arrangement with an arc and a counter-image on the
  other side, but the spaghetti diagram is completely asymmetric.  LQ
  can be visually distinguished from SQ from the relative distances of
  the arc and the counter-image.  For LQ the arc is closer, for SQ the
  counter-image is closer.
\end{itemize}
Although the labels `short-axis quad' and so on are not standard in
the literature, the morphological classification they express is
familiar to experienced modellers.  Hence they can be useful to
researchers wishing to apply other modelling methods to the same
systems.

>>>>

We now discuss the diagnostics that can be given from the process of
drawing the spaghetti diagram, even before detailed mass-modelling
takes place.  Figure~\ref{fig:splinput} shows nine examples, each
consisting of a cutout of the Space~Warps image of a lens candidate,
marked up with a spaghetti diagram.

@@

The fourth diagnostic is based on the fact that the arrangement of
lensed images of a pointlike source through a non-circular
gravitational lens depends on the location of the source relative to
the long and short axes of the lens.  This dependence is quite robust
and independent of many other details of the lensing mass
distribution.  Sources close to being dead-centre behind a lens tend
to produce quads, sources at larger transverse distance tend to
produce doubles.  We denote these as Q and D respectively, and add a
prefix to the Q systems, as follows: we write LQ if the source is
inferred as displaced along the long axis of the lens, SQ if displaced
along the short axis, IQ if inclined to both axes, CQ if only very little to
no inclination.  Although the
unlensed source and its location are obviously not seen, the LQ, SQ, IQ and CQ
cases correspond to easily-seen image morphologies \citep[see,
  e.g.,][]{2003AJ....125.2769S}.
\begin{itemize}
\item The simplest is the SQ case: this creates a saddle point and two
  minima in an arc, with the second saddle point on the other side of
  the lensing galaxy, and closer to the galaxy than the arc is.  SW58
  and SW28 in the upper row of Figure~\ref{fig:splinput} are typical
  examples of SQ.
\item The middle row of Figure~\ref{fig:splinput} shows three IQ
  systems, SW05, SW42 and SW19.  This type has a characteristic
  asymmetry, often with two images close together.
\item The lower row of Figure~\ref{fig:splinput} shows three LQ
  systems, SW09, SW29 and SW02, and the failed model SW57 at top right
  also possibly belonging to this category.  Here the images have a
  fairly symmetric arrangement with an arc and a counter-image on the
  other side, but the spaghetti diagram is completely asymmetric.  LQ
  can be visually distinguished from SQ from the relative distances of
  the arc and the counter-image.  For LQ the arc is closer, for SQ the
  counter-image is closer.
\end{itemize}
Although the labels `short-axis quad' and so on are not standard in
the literature, the morphological classification they express is
familiar to experienced modellers.  Hence they can be useful to
researchers wishing to apply other modelling methods to the same
systems.
